Deep Reinforcement Learning for Conservation Decisions

### 2021-05-18 manuscript/manuscript.Rmd
Small changes to formal section

The reinforcement learning problem is typically formalized as a discrete-time partially observable Markov decision process (POMDP).
A POMDP is a 7-tuple that consists of the following: 
- $\mathcal{S}$: a set of states called the state space
- $\mathcal{A}$: a set of actions called the action space
- $\Omega$ : a set of observations called the  observation space
- $E(o_{t}| s_{t})$: an emission distribution, which accounts for an agent's observation being different from environment's state
- $T(s_{t+1}|s_t, a_t)$: a state transition operator which describes the dynamics of the system
- $r(s_t, a_t)$: a reward function
- $d_0(s_0)$: an initial state distribution
- $\gamma \in (0,1]$: a discount factor

After the agent makes an observation at time t, $o_t$, the agent selects an action, $a_t$, according to its policy, $\pi$, which maps observations to a probability distribution over actions, $\pi: \Omega \rightarrow p(\mathcal{A} \mid \Omega)$.
This action will cause the environment to transition to a new state according to its state transition operator. 
The agent will then receive a new observation and the process will repeat until a terminal state is reached.
As the agent interacts with the environment, the agent will create a trajectory in state-action space given as $\tau = (s_0, a_0, \dots, s_H, a_H)$ where H denotes the horizon or length of the state-action sequence.
From these definitions, we can provide an agent's trajectory distribution for a given policy as,

@@

An important property of POMDPs and MDPs is that they are memoryless; the future, $s_{t+1}$, does not depend on the past but only on the state and action at the present time.
This Markov property, which is often an unrealistic assumption, is held in most of the RL algorithms that we discuss in the next section.

In formal terms, the goal of reinforcement learning is to find an optimal policy distribution, $\pi^*(a_t|o_t)$, that maximizes the expected return, $J(\pi)$:

@@

Note that this formalism extends to an infinite horizon case when $T= \infty$.
While there are RL methods to deal with infinite horizon cases, we will only present methods for finite horizon or episodic POMDPs.

>>>>

In this brief section, we will introduce the concept of the RL environment as a POMDP; we will then use POMDP definitions to ultimately define the RL objective.
Since all RL algorithms are either directly or indirectly maximizing the RL objective, understanding the RL objective is crucial to understanding how RL algorithms work. <!-- ML: this seems like an obvious sentence to me but am stuck on it atm so revisions welcomed-->

The reinforcement learning environment is typically formalized as a discrete-time partially observable Markov decision process (POMDP).
A POMDP is a 7-tuple that consists of the following: 
- $\mathcal{S}$: a set of states called the state space
- $\mathcal{A}$: a set of actions called the action space
- $\Omega$ : a set of observations called the observation space
- $E(o_{t}| s_{t})$: an emission distribution, which accounts for an agent's observation being different from environment's state
- $T(s_{t+1}|s_t, a_t)$: a state transition operator which describes the dynamics of the system
- $r(s_t, a_t)$: a reward function
- $d_0(s_0)$: an initial state distribution
- $\gamma \in (0,1]$: a discount factor

The agent interacts with the environment in an iterative loop, whereby the agent only has access to the observation space, action space and the discounted reward signal, $\gamma^t \, r(s_t, a_t)$.
At each time step, the agent makes an observation of the environment at time t, $o_t$, and selects an action, $a_t$, according to its policy, $\pi$, which maps observations to a probability distribution over actions, $\pi: \Omega \rightarrow p(\mathcal{A} \mid \Omega)$.
The action will cause the environment to transition to a new state according to the state transition operator. 
The agent will then receive a new observation and a discounted reward signal, $\gamma^t \, r(s_t, a_t)$, as feedback. 
This process will repeat until a terminal state is reached.
As the agent interacts with the environment, the agent will create a trajectory in state-action space given as $\tau = (s_0, a_0, \dots, s_H, a_H)$ where H denotes the horizon or length of the state-action sequence.
From these definitions, we can provide an agent's trajectory distribution for a given policy as,

@@

The goal of reinforcement learning is for the agent to find an optimal policy distribution, $\pi^*(a_t|o_t)$, that maximizes the expected return, $J(\pi)$:

@@

Note that this formalism extends to an infinite horizon case when $T= \infty$.
While there are RL methods to deal with infinite horizon cases, we will only present methods for finite horizon or episodic POMDPs.
In the appendix, we will go into further detail on the general approaches that model-free deep RL algorithms take towards optimizing the RL objective.
